---
title: "Project Draft"
output: html_document
date: "2023-11-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
source("../scripts/viridis.R")
source("../scripts/ggprob.R")
```

# Group members
* Hareem 
* Vishnu
* Sanjay
* Sumin 

# Introduction
    
* For our analysis, we decided to look into the stock prices in earning and non-earning seasons from May 17, 2012 to August 02, 2020. The earning period includes January, April, July, October. The non-earning period includes February, March, May, and November. We observed this specific time period due to the fact that this range contains the complete data for all five FAANG companies we want to observe in our analysis. The motivation behind analyzing stock prices in this way was to obtain information on market conditions, but more specifically to observe the differences in fluctuations between each FAANG company. Our question of interest states: During earnings seasons (e.g., January, April, July, October), do the stock prices of these companies experience larger fluctuations compared to non-earnings periods (e.g., February, March, May, November). Our study aims to determine if earnings seasons, compared to non-earnings periods, significantly affect stock price volatility for FAANG companies from 2012 to 2020, utilizing daily percentage changes in opening prices and the Difference in Means methodology.


# Background

* Describe the data set, how it was collected, what the variables mean.
  * The data set was collected through web scraping of financial data from Yahoo Finance. This web scraping process involved extracting information related to the stock prices of the five prominent companies known as FAANG, which includes Facebook, Amazon, Apple, Netflix, and Google. Each case in the data set represents a specific date or trading day. The data set likely includes multiple rows, each corresponding to a different trading day, for each of the five FAANG companies. Each row contains various information about the stock prices and trading volumes for each trading day from 1980s - 2020s (depending on the company), we will be looking at the data set from 2012-2020 since that is the earliest available data for all five companies.
  
  * Key Variables:
    * Opening Price: The price of a stock at the beginning of the trading day.
    * Closing Price: The price of a stock at the end of the trading day.
    * High Price: The highest price reached by the stock during the trading day.
    * Low Price: The lowest price reached by the stock during the trading day.
    * Date: The date of the trading day for which the stock price data is recorded.
    
  * Data Size: The data set consists of multiple CSV files, with each file corresponding to one of the five FAANG companies. Each CSV file contains historical stock price data, typically with multiple rows, where each row represents a trading day. The total size of the data set would depend on the time period covered but could include several years of daily trading data for each company.
  
* Describe any background information needed to better comprehend the question or questions you are posing and how the data relate to the question.
  * To fully grasp the question we are posing in this project, we have to understand what an earning period is. An earning period is the time when public companies release their quarterly earnings report. This new piece of information is important because it can create volatility in the market. Another clarification to make is how we define “large fluctuations”. We will determine this by looking at the magnitude of change between the stock prices of the earning period vs the non earning period. In our question we explicitly state that we will be analyzing data from the five FAANG companies. The reason we narrowed it down to only those five was because these companies are the largest and most successful globally, and we were able to find solid data from these companies to use in our analysis. Lastly, we chose the time period of May 17, 2012 to August 02, 2020 because in order to effectively compare the five FAANG companies, we had to make sure we have complete data from all five companies.

  
* Describe any unusual factors which may affect interpretation of results.
  *There are several outlying factors that could influence our results but the four that have probably been the most obvious and well-known factors during recent times are the 2016 and 2020 Presidential elections, COVID-19, global political and economic conditions, and each of the FAANG companies' individual policies have all influenced their stock market trends in one way or another. Hence, it’s impossible to say that our results would be 100% accurate.

* Describe what you intend to do in the rest of the report.
  * In the report's next phases, we will analyze stock price changes for FAANG companies from 2012 to 2020. Using graphs and stats like daily stock prices and market shifts, we will compare between earnings and non-earnings seasons. Using the information we find during these periods, we will create a statistical summary regarding our findings, considering any unusual circumstances in the data and suggesting ways to improve it.
  

# Analysis

>We first read stock price data from our company csv files, and only include data from 2012-5-18 to 2020-08-18 (This is the earliest and latest dates for which we have readings from every company). We alter the data to include a column for whether they in an Earning Period or not depending on the months of the year, and create summary tables that group by year and period and calculate the average absolute percentage change from Open price to Close price for that period.
The first few rows of these summaries for the companies are as follows (Every company has a summary table in the same format:

```{r, include = FALSE}
Amazon_data = read_csv("../data/Amazon.csv")
tidy_Amazon = Amazon_data %>%
  filter(between(Date, as.Date("2012-05-18"), as.Date("2020-08-18")))
```

```{r, include = FALSE}
Apple_data = read_csv("../data/Apple.csv")
tidy_Apple = Apple_data %>%
  filter(between(Date, as.Date("2012-05-18"), as.Date("2020-08-18")))
```

```{r, include = FALSE}
Facebook_data = read_csv("../data/Facebook.csv")
tidy_Facebook = Facebook_data %>%
  filter(between(Date, as.Date("2012-05-18"), as.Date("2020-08-18")))
```

```{r, include = FALSE}
Google_data = read_csv("../data/Facebook.csv")
tidy_Google = Google_data %>%
  filter(between(Date, as.Date("2012-05-18"), as.Date("2020-08-18")))
```

```{r, include = FALSE}
Netflix_data = read_csv("../data/Netflix.csv")
tidy_Netflix = Netflix_data %>%
  filter(between(Date, as.Date("2012-05-18"), as.Date("2020-08-18")))
```

## Average Absolute Percentage Change in Amazon Stock Prices: Earning vs. Non-Earning Periods
> Here, we are conducting a difference in the absolute mean percentage change in stock prices for Amazon during different periods of the year - Earning Periods (EP) and Non-Earning Periods (NEP).

### For each table, the columns are as follows: 

* Year: Year for which we look at a specific period
* Period: Which period we are considering. EP represents Earnings Period (January, April, July, October), and NEP represents Non Earnings Period (February, March, May, June, August, September, November, December)
* Percentage_Change: The average percentage change for that period and year, for which an individual day percentage change is calculated by the absolute difference between opening and closing price for a day divided by the opening price in percent.

```{r, include = FALSE}
tidy_Amazon <- tidy_Amazon %>%
  mutate(Period = case_when(
    month(Date) %in% c(1, 4, 7, 10) ~ "EP",
    month(Date) %in% c(2, 3, 5, 6, 8, 9, 11, 12) ~ "NEP"
  ))

# Summary for Amazon
summary_Amazon <- tidy_Amazon %>%
  # Filter for relevant months
  filter(Period %in% c("EP", "NEP")) %>%
  # Group by Year and Period
  group_by(Year = year(Date), Period) %>%
  # Calculate the percentage change
  summarize(Percentage_Change = abs(mean((Close - Open) / Open * 100))) %>%
  ungroup()

summary_Amazon
```

```{r, include = FALSE}
tidy_Apple <- tidy_Apple %>%
  mutate(Period = case_when(
    month(Date) %in% c(1, 4, 7, 10) ~ "EP",
    month(Date) %in% c(2, 3, 5, 6, 8, 9, 11, 12) ~ "NEP"
  ))

# Summary for Amazon
summary_Apple <- tidy_Apple %>%
  # Filter for relevant months
  filter(Period %in% c("EP", "NEP")) %>%
  # Group by Year and Period
  group_by(Year = year(Date), Period) %>%
  # Calculate the percentage change
  summarize(Percentage_Change = abs(mean((Close - Open) / Open * 100))) %>%
  ungroup()

summary_Apple
```

```{r, include = FALSE}
tidy_Facebook <- tidy_Facebook %>%
  mutate(Period = case_when(
    month(Date) %in% c(1, 4, 7, 10) ~ "EP",
    month(Date) %in% c(2, 3, 5, 6, 8, 9, 11, 12) ~ "NEP"
  ))

# Summary for Amazon
summary_Facebook <- tidy_Facebook %>%
  # Filter for relevant months
  filter(Period %in% c("EP", "NEP")) %>%
  # Group by Year and Period
  group_by(Year = year(Date), Period) %>%
  # Calculate the percentage change
  summarize(Percentage_Change = abs(mean((Close - Open) / Open * 100))) %>%
  ungroup()

summary_Facebook
```

```{r, include = FALSE}
tidy_Google <- tidy_Google %>%
  mutate(Period = case_when(
    month(Date) %in% c(1, 4, 7, 10) ~ "EP",
    month(Date) %in% c(2, 3, 5, 6, 8, 9, 11, 12) ~ "NEP"
  ))

# Summary for Amazon
summary_Google <- tidy_Google %>%
  # Filter for relevant months
  filter(Period %in% c("EP", "NEP")) %>%
  # Group by Year and Period
  group_by(Year = year(Date), Period) %>%
  # Calculate the percentage change
  summarize(Percentage_Change = abs(mean((Close - Open) / Open * 100))) %>%
  ungroup()

summary_Google
```

```{r, include = FALSE}
tidy_Netflix <- tidy_Netflix %>%
  mutate(Period = case_when(
    month(Date) %in% c(1, 4, 7, 10) ~ "EP",
    month(Date) %in% c(2, 3, 5, 6, 8, 9, 11, 12) ~ "NEP"
  ))

# Summary for Amazon
summary_Netflix <- tidy_Netflix %>%
  # Filter for relevant months
  filter(Period %in% c("EP", "NEP")) %>%
  # Group by Year and Period
  group_by(Year = year(Date), Period) %>%
  # Calculate the percentage change
  summarize(Percentage_Change = abs(mean((Close - Open) / Open * 100))) %>%
  ungroup()

summary_Netflix
```

```{r, include = FALSE, echo = FALSE}
print("Amazon")
print(head(summary_Amazon,5))
print("Apple")
print(head(summary_Apple,5))
print("Google")
print(head(summary_Google,5))
print("Netflix")
print(head(summary_Netflix,5))
print("Facebook")
print(head(summary_Facebook,5))
```

## Graph
> The graphical summary is a boxplot created using ggplot2 in R. This plot visually represents the distribution of percentage changes in Amazon's stock prices across different years and for both EP and NEP.

### Key aspects of the boxplot:

* X-axis: Represents the year.
* Y-axis: Shows the absolute mean percentage change in stock price.
* Boxplot Structure: Each box shows the median, quartiles, and outliers in the percentage change data for each year and period.
* Color Coding: Different colors for EP and NEP (blue for EP, red for NEP) allow for quick visual distinction.
This boxplot is particularly effective in showing how the variability in stock price changes differs between earning and non-earning periods over the years. It allows for a quick visual comparison of trends and variances across different periods.

## Inferences and Evidence
> From the numerical and graphical summaries, several inferences can be made:

* Variability Over Time: By examining the spread and central tendency in the boxplot, we can infer if there's more variability in stock prices during earning or non-earning periods.
* Yearly Trends: The plot can indicate whether there's a trend in stock price changes over the years.
* Each of these inferences is backed by the data presented in the summaries. The numerical summary provides exact figures, while the graphical summary offers a visual representation that can sometimes reveal patterns not immediately obvious in the numerical data.


Amazon:
```{r, echo = FALSE}
ggplot(summary_Amazon, aes(x = Period, y = Percentage_Change, fill = Period)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Amazon Stock Price Change Boxplot",
       x = "Earning vs. Non-Earning Periods",
       y = "Percentage Change") +
  scale_fill_manual(values = c("EP" = "blue", "NEP" = "red"))
```

Apple:
```{r, echo = FALSE}
ggplot(summary_Apple, aes(x = Period, y = Percentage_Change, fill = Period)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Apple Stock Price Change Boxplot",
       x = "Earning vs. Non-Earning Periods",
       y = "Percentage Change") +
  scale_fill_manual(values = c("EP" = "blue", "NEP" = "red"))
```

Facebook:
```{r, echo = FALSE}
ggplot(summary_Facebook, aes(x = Period, y = Percentage_Change, fill = Period)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Facebook Stock Price Change Boxplot",
       x = "Earning vs. Non-Earning Periods",
       y = "Percentage Change") +
  scale_fill_manual(values = c("EP" = "blue", "NEP" = "red"))
```

Google:
```{r, echo = FALSE}
ggplot(summary_Google, aes(x = Period, y = Percentage_Change, fill = Period)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Google Stock Price Change Boxplot",
       x = "Earning vs. Non-Earning Periods",
       y = "Percentage Change") +
  scale_fill_manual(values = c("EP" = "blue", "NEP" = "red"))
```

Netflix:
```{r, echo = FALSE}
ggplot(summary_Netflix, aes(x = Period, y = Percentage_Change, fill = Period)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Netflix Stock Price Change Boxplot",
       x = "Earning vs. Non-Earning Periods",
       y = "Percentage Change") +
  scale_fill_manual(values = c("EP" = "blue", "NEP" = "red"))
```

> We will now conduct a difference of means t test between the Earning Periods and Non Earning Periods values for each company to see if there is a significant likelihood that their means vary.
The Null Hypothesis is that there is no significant difference between the means of the two groups for a given company.
The alternative Hypothesis is that there is a significant difference between the means of the two groups for a given company.

```{r, include = FALSE, echo = FALSE}

company_names <- c("Amazon", "Facebook", "Google", "Netflix", "Apple")

# Loop through each company
for (company in company_names) {
  # Create the table name dynamically
  table_name <- paste("summary_", company, sep = "")
  
  # Create the pivot table name dynamically
  pivot_table_name <- paste("pivot_", company, sep = "")
  
  # Pivot the table
  assign(pivot_table_name, get(table_name) %>%
           pivot_wider(names_from = Period, values_from = Percentage_Change))
}



pivot_names <- c("Amazon", "Facebook", "Google", "Netflix", "Apple")

# Perform t-test for each pivot table
for (pivot_name in pivot_names) {
  # Get the pivot table dynamically
  pivot_table <- get(paste0("pivot_", pivot_name))
  
  # Perform t-test on EP and NEP columns
  t_test_result <- t.test(pivot_table$EP, pivot_table$NEP)
  
  # Print the results
  cat("Company:", pivot_name, "\n")
  cat("t-test p-value:", t_test_result$p.value, "\n")
  cat("\n")
}

```

> All the p-values we have obtained from the companies are above 0.05

> To get some more information, let us plot these percentages, EP vs NEP, year by year.

## Line Graphs:
* Trend Analysis: Line graphs are excellent for showing trends over time. They clearly depict how the average stock price fluctuation (either during earning or non-earning periods) changes from year to year.
* Comparative Analysis: By plotting lines for both earnings and non-earnings months on the same graph, we can directly compare these two trends year by year. It becomes easier to see whether fluctuations are generally greater during one period or the other across different years.
* Smoothness and Continuity: Line graphs provide a sense of continuity, which is particularly useful for financial data. They help in visualizing the progression and direction of changes over time in a way that box plots, which focus on distribution and variability.
* Immediate Visual Impact: Line graphs can sometimes convey information more immediately and are often easier for a general audience to understand at a glance. They can highlight specific points of interest, like spikes or drops in stock price fluctuations.

### Summary

* The box plots are useful for examining the distribution of stock price changes for each year and identifying any outliers or significant variations within a year.
* The line graphs complement this by showing how the average fluctuation changes year over year, providing a clear picture of upward or downward trends.
Conclusion:


```{r}
# Function to create and display the plot
create_and_display_plot <- function(pivot_table, company_name) {
  avg_fluctuation <- pivot_table %>%
    group_by(Year) %>%
    summarise(
      Avg_Fluctuation_Earnings = mean(EP),
      Avg_Fluctuation_NonEarnings = mean(NEP)
    )
  
  ggplot(avg_fluctuation, aes(x = Year)) +
    geom_line(aes(y = Avg_Fluctuation_Earnings, color = "Earnings Months")) +
    geom_line(aes(y = Avg_Fluctuation_NonEarnings, color = "Non-Earnings Months")) +
    labs(title = paste("Average Fluctuation in Stock Prices Over Time -", company_name),
         x = "Year",
         y = "Fluctuation in Stock Prices",
         color = "Legend") +
    scale_color_manual(values = c("Earnings Months" = "blue", "Non-Earnings Months" = "red")) +
    theme_minimal()
}

# Create and display plots for each company
plot_Amazon <- create_and_display_plot(pivot_Amazon, "Amazon")
print(plot_Amazon)

plot_Facebook <- create_and_display_plot(pivot_Facebook, "Facebook")
print(plot_Facebook)

plot_Google <- create_and_display_plot(pivot_Google, "Google")
print(plot_Google)

plot_Netflix <- create_and_display_plot(pivot_Netflix, "Netflix")
print(plot_Netflix)

plot_Apple <- create_and_display_plot(pivot_Apple, "Apple")
print(plot_Apple)

```


> The provided code generates line plots for the average fluctuation in stock prices over the years 2012 to 2020 for the five FAANG companies (Amazon, Facebook, Google, Netflix, and Apple). The average fluctuation is calculated separately for earnings months and non-earnings months, where earnings months typically include periods such as January, April, July, and October, and non-earnings months include months like February, March, May, and November.

> The purpose of these plots is to visually inspect and compare the trends in stock price fluctuation between earnings and non-earnings months for each FAANG company over the specified time period. By plotting these lines on the same graph, it allows for a direct visual comparison between the two types of months for each company.

### Discussion
* Provide broader interpretations of your analysis and describe how to interpret your results with respect to your question of interest.
  * Our analysis explores whether earnings seasons yielded a substantial influence on stock price fluctuations for FAANG companies from 2012 to 2020 over non-earning seasons. The statistical summary, when compared between earning and non-earning periods, aims to showcase whether there's a notable difference in stock price movements during these distinct times. If the graphs demonstrate greater volatility (rapid/unpredictable changes) in stock prices during earnings seasons, it implies a potential correlation between earning announcements and stock market reactions. However, these findings should be considered in addition to the stock market context and external factors that influenced stock price that might not be captured within our data set during our selected time frame.

* Discuss any potential short-comings of the analysis.
  * While we have tried our best to show the relationship between earring seasons and stock price volatility, it's necessary to acknowledge the barriers that we had to confront. The project primarily focuses on earnings-related fluctuations, but it might overlook other crucial factors influencing stock prices, such as broader economic trends, market reactions, or unexpected company-specific developments. Furthermore, our ability to attribute all fluctuations solely to earning and non-earning seasons might be constrained. Additionally, the data set's time frame does not consider all relevant market dynamics, potentially limiting the comprehensiveness of our conclusions. Understanding these limitations is essential in order to interpret the results we discover.
  
* Discuss potential future directions for additional work
  * In this project, we analyzed stock price fluctuations in earning vs non-earning periods. This question effectively sets the basis to explore additional questions more in depth. In the future, when we have records of more data, we can observe the stock prices a few years prior to, during, and years post the Coronavirus pandemic. This data would be interesting to observe in terms of unemployment rates because of how drastically unemployment changed during these time periods and how earning periods were impacted by these changes. An additional question we can explore is: Are there significant differences in the stock price fluctuations of FAANG companies (Facebook, Amazon, Apple, Netflix, and Google) during their earnings seasons? This question would help us view the differences between the companies and find the causes for these differences. An important factor that could be driving these differences is each company's business models and plans. An alternate method we can use to address this same question is make a bar plot for each year in which we look at the fluctuations by month to see if there is a specific pattern certain months of either period follow across the years. For example, this will give us a closer look to see if the stock prices in January (the first month of the earning period) is significantly different than October (the last month of the earning period). This method will allow us to observe patterns within the two periods in addition to comparing them to each other. The most helpful data we can collect to refine our understanding would definitely be stock price fluctuations post Coronavirus. This would truly provide us more insight on if and how stock prices are impacted by economic depressions. In addition, we could also collect data regarding GDP growth and compare that with stock fluctuations to see if overall the stock prices were high during a time of increased GDP growth. 
  
* Summarize your primary conclusions and the primary evidence that supports these conclusions.
  * The primary conclusion we can draw from our analysis is that overall, there is no significant difference between the earning vs non-earning period for each company. Looking directly at the box plot, we can see that the difference between the median percent change values of the earning and non-earning periods isn't very significant. This can be better explained by comparing the means of the two periods where the p-values are at a range that accept the null hypothesis across all the companies. Since the p-values of all five FAANG companies are above 0.05 (the significance level), we don’t have strong statistical evidence to reject the null hypothesis that there is no significant difference between the means of the two groups for a given company. Additionally, if we look at the average fluctuation in prices by year in the line graphs, we can see that the differences of the percentages of the earning period vs non-earning period isn't significantly different either. The actual percentage of earning period and non-earning period is different since we are comparing different months of the year, but we are concerned with the range. The differences of the percentages are within the same range. 

### References
* Cite the source of your data.
  * https://www.kaggle.com/datasets/aayushmishra1512/faang-complete-stock-data
  * ** The information in the kaggle data set is from Yahoo Finance. 
  * Collection Methodology: Web Scraping

